The Galactic Midplane is Not a Plane: Implications for dynamical analysis with {\em Gaia} data and beyond

### 2019-04-22 paper/ms.tex
Change to J_R, J_phi throughout

Once a Galactocentric coordinate system has been established and a
six-dimensional (6D) phase space measurement of a star has been made, it is
often desirable to convert this measurement into action space to concisely
summarize its projected orbit, model the stellar distribution function, or
find stars with similar dynamical properties. Actions are conserved quantities
that describe the orbit of a star under the assumption of regular, bound
orbits in a system where the equations of motion are separable in a particular
coordinate system. They are the cyclical integral of the canonical momentum
over its conjugate position,
\beq\label{eq:actions}
J_i \equiv
\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,.}
\eeq
Under the assumption of axisymmetry, $i=R,z,\phi$ are the radial, vertical,
and azimuthal coordinates respectively in a cylindrical coordinate system and
the $p_i$ are the conjugate momenta.\footnote{We write $J_r$ instead of $J_R$
for aesthetic reasons.} In a slowly-evolving axisymmetric potential, these
actions are invariant and $J_{\phi} \equiv L_z$
\citep{2008gady.book.....B,2014RvMP...86....1S}.

@@

\begin{deluxetable*}{cCCCCCCCC}
\tablecaption{Relevant dynamical parameters for the $9$ open clusters within
$250\,\pc$ of the Sun, using the relevant initial positions and velocities
from \citet{2018AA...616A..10G}. The neighborhood column is explained in more
detail in Section~\ref{sssec:reconstruction}.\label{tab:real_clusters}}
\tablehead{\colhead{cluster} & \colhead{$N_{\text{memb}}$} & \colhead{distance} & \colhead{$J_r$} &
\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} & \colhead{mass} & \colhead{neighborhood} \\ 
\colhead{ } & \colhead{ } &
\colhead{$\mathrm{pc}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{pc}$} & \colhead{$M_{\odot}$} & \colhead{$\mathrm{pc}$}}
\startdata
alphaPer & 740 & 174.9 & \sim20 & \sim -1700 & \sim0.005 & \sim11 & 352\tablenotemark{b} & 57 \\
Blanco1 & 489 & 237.2 & \sim2 & \sim-1800 & \sim2 & \sim210 & 410\tablenotemark{c} & 1460 \\
ComaBer & 153 & 85.9 & \sim2 & \sim-1900 & \sim0.7 & \sim140 & 112\tablenotemark{d} & 522\\
Hyades & 515 & 47.5 & \sim20 & \sim-1800 & \sim0.3 & \sim80 & 400\tablenotemark{e} & 462 \\
IC2391 & 325 & 151.6 & \sim6 & \sim-1800 & \sim0.02 & \sim20 & \nodata & \nodata \\
IC2602 & 492 & 152.2 & \sim10 & \sim-1700 & \sim0.3 & \sim89 & \nodata & \nodata \\
NGC2451A\tablenotemark{a} & 400 & 193.7 & \sim6 & \sim-1800 & \sim0.2 & \sim73 & \nodata & \nodata \\
Pleiades & 1326 & 135.8 & \sim20 & \sim-1700 & \sim0.4 & \sim96 & 800\tablenotemark{f} & 685 \\
Praesepe & 938 & 186.2 & \sim20 & \sim-1800 & \sim0.6 & \sim120 & 550\tablenotemark{d} & 781
\enddata

@@

We can write down the actions in this approximation as
\citep[Section~3.5.3b][]{2008gady.book.....B},
\beq\label{eq:actions_epi}
\begin{split}
L_z &= R_g v_c \\
J_r &= \kappa A_R^2 \\
J_z &= \nu A_z^2\text{,}
\end{split}
\eeq
where $R_g$ is the guiding radius of the orbit.

An induced error in $z$ of $\Delta z$ will cause $\Delta A_z \sim \Delta z$.
Similarly, an induced error in $R$ will cause $\Delta A_R \sim \Delta R$.
Making the assumption that $\kappa$ and $\nu$ are constant, we have that
\beq\label{eq:actions_epi_error}
\begin{split}
\Delta L_z = \Delta R v_c &\implies \frac{\Delta L_z}{L_z} = \frac{\Delta R}{R} \\
\Delta J_r = 2 \kappa \Delta A_R &\implies \frac{\Delta J_r}{J_r} = \frac{2\Delta R}{A_R} \\
\Delta J_z = 2 \nu \Delta A_z &\implies \frac{\Delta J_z}{J_z} = \frac{2\Delta z}{A_z}\text{.}
\end{split}
\eeq
Therefore, errors in each action are linearly proportional to the errors in
either the midplane or cylindrical radius. While in detail $\kappa$ and $\nu$
are not constant, we will find that this order of magnitude estimate matches
the numerical estimates quite well.

There is a slight complication in the formula for computing $\Delta J_r/J_r$
and $\Delta L_z/L_z$. In a moment we will consider the effect of an inaccuracy
in the distance from the Sun to the Galactic center on actions. However, this
inaccuracy induces an error in $x$, or $\Delta x$. This is not the same as an
inaccuracy in $\Delta R$, but they are related by the following argument. An
error $\Delta x$ will induce an erroneous radius $R_{\text{err}}$ related by
the formula,
\beq
(x+\Delta x)^2 + y^2 = R_{\text{err}}^2\text{.}
\eeq
Keeping only terms to first order in $\Delta x$, we have that,
\beq
\begin{split}
R_{\text{err}}^2 &= R^2 - 2 R \cos{\theta} \Delta x \\
\implies \Delta R &\equiv \abs{R_{\text{err}} - R} = \abs{\cos{\theta}} \Delta x\text{.}
\end{split}
\eeq
Averaging over the circle, we therefore have that
\beq
\avg{\Delta R} = \frac{2}{\pi} \Delta x\text{.}
\eeq
Therefore, the error in $J_r$ and $L_z$ are related to $\Delta x$ by,
\beq\label{eq:Jr_epi_dx}
\begin{split}
\frac{\Delta J_r}{J_r} &= \frac{4 \Delta x}{\pi A_R} \\
\frac{\Delta L_z}{L_z} &= \frac{2 \Delta x}{\pi R}\text{.}
\end{split}
\eeq

A similarly rough argument can be made to arrive at the following formulae for
the impact of velocity offsets on the actions,
\beq\label{eq:actions_epi_error}
\begin{split}
\frac{\Delta L_z}{L_z} &= \frac{\Delta v_{\phi}}{v_c} \\
\frac{\Delta J_r}{J_r} &= \frac{\Delta v_R}{v_{R,\text{max}}} \\
\frac{\Delta J_z}{J_z} &= \frac{\Delta v_z}{v_{z,\text{max}}}\text{,}
\end{split}
\eeq
where $v_{R,\text{max}}$ and $v_{z,\text{max}}$ are the maximum radial and
vertical velocities of the orbit, respectively.

@@

\begin{deluxetable*}{cccccccc}
\tablecaption{Description and names of the three orbits considered in this
work. For the last two columns: $z_{\text{max}}$ the maximum height of the
orbit while $A_R$ is the magnitude of the radial excursions of the orbit,
computed as one half the maximum radius minus the minimum
radius.\label{tab:orbits}}
\tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
velocity} & \colhead{$J_r$} &
\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
\colhead{$A_R$}\\ \colhead{ } &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$}}
\startdata 
thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 & 1.29 \\
thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 & 1.19 \\ 
halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16 & 2.34
\enddata
\end{deluxetable*}

@@

For a systematic offset in $z$ ({\em upper} panels), the
95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
$6.2\,\actunit$ for $J_r$ and $J_z$, respectively. These are fractional errors
of $5.7\,\%$ and $85.7\,\%$, respectively. The error induced in $L_z$ is
negligible, as expected since $L_z$ is independent of $z$. It is worth
pointing out that a $100\,\pc$ error in an orbit with
$z_{\text{max}}=850\,\pc$ --- a $12\%$ error
--- induced an $86\%$ error in the computation of $J_z$.

For a systematic offset in $x$ (or Galactic center distance), the
95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$,
$47$, and $0.71\,\actunit$ for $J_r$, $L_z$ and $J_z$, respectively. These are
fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively.

\begin{figure*}
\plotone{fig/schmactions_one_orbit.pdf}
\caption{The artificial phase-dependence in the observed actions induced by an
error in the Galactocentric coordinate system. We consider here the thick-disk
orbit, which has actions of $(J_r, L_z, J_z) = (37.9, -1520, 7.0)\,\actunit$
and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We integrate the
orbit according to the procedure laid out in Section~\ref{ssec:action_comp}.
Then, we subtract $100\,\pc$ from the $z$ value ({\em upper panels}) or the
$x$ value ({\em lower panels}) of each position in the orbit, corresponding to
an erroneous observer assuming a midplane ({\em upper}) or solar radius ({\em
lower}) that is off by $100\,\pc$. We then allow our (immortal) observer to
observe the orbit over $1\,\Gyr$ and perform the same orbit integration
procedure at each timestep, and report the values of the actions. The
computation of $L_z$ is pristine to errors in $z$, with only numerical
artifacts remaining. Only small errors are induced in $J_r$, with the middle
$90\%$ of values over the $\Gyr$ being within $\sim8\%$ of the true $J_r$. As
expected, large errors are induced in $J_z$ with a $100\,\pc$ offset in $z$,
with the middle $90\%$ of values being within $\sim43\%$ of the true $J_z$.
The $x$ offset induces uncertainties in $J_r$ and $L_z$ of $\sim21\%$ and
$\sim3\%$, respectively. A $\sim3\%$ error in $J_z$ is also induced.}
\label{fig:one_orbit_wrong_ref}
\end{figure*}

\begin{figure}
\plotone{fig/schmactions_Jz_hist.pdf}
\caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
values have a bias (from e.g. an incorrectly computed midplane), then the
induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
sort of error propagation must take this into account. Intuition for the shape
of each panel is given in the text. We also plot one half the 95\uth{}
percentile minus the 5\uth{} percentile of each distribution as a horizontal
arrow anchored on the true $J_z$ value. We call this $\Delta J_z$ and will use
it (along with the similarly defined $\Delta J_r$ and $\Delta L_z$) to
empirically describe the error distribution. We see that $\Delta J_z$ roughly
corresponds to the distance from the true $J_z$ value to one of the modes of
the error distribution.}
\label{fig:Jz_hist}
\end{figure}

@@

The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
in each action for an offset in the $z$-component (i.e. the Galactic
midplane). In the {\em bottom row} we consider offsets in the $x$ component
(i.e. the Solar radius). The {\em left}, {\em middle}, and {\em right} columns
show the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$,
respectively. We compute the fractional spread $\Delta J_i/J_i$ as one half
the 95\uth{} minus 5\uth{} percentile of the action computed using the offset
coordinate system over the course of the first $\Gyr$, divided by the true
action (computed in the correct coordinate system).

In the {\em upper middle} panel of Figure~\ref{fig:many_orbit_wrong_ref},
there is essentially no spread in the determination of $L_z$. This is expected
since, in principle, $L_z$ only depends on the $x$- and $y$-components of the
position and velocity of the stars\footnote{In practice, however, $L_z$ is
computed as part of the torus-fitting method.}, and is thus unaffected by
offsets in $z$. Indeed, the result we found in
Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all
orbit types.

@@

For the offset in the Solar radius ({\em bottom panels}), the error is largest
for $J_r$, with some deviations resulting in $L_z$ and relatively small
deviations in $J_z$. In the {\em bottom middle} and {\em bottom right} panels
all three lines nearly overlap.

\begin{figure*}
\plotone{fig/schmactions_many_orbits.pdf}
\caption{We report one half the 95\uth{} minus 5\uth{} percentile of the error
in the measured action ($\Delta J_i$) from coordinate system errors for the
thin, thick, and halo orbits (Table~\ref{tab:orbits}). See discussion in the
text for the justification in using this to measure the magnitude of the
induced error. The {\em left}, {\em center}, and {\em right} panels show the
result for $J_r$, $L_z$, and $J_z$, respectively. The {\em upper} panels
consider an offset in $z$ and the {\em lower} panels consider an offset in $x$
(equivalently, an offset in the Solar radius). In some panels we also plot as
dashed lines the epicyclic prediction of the induced action error
(Equations~\ref{eq:actions_epi_error}~and~\ref{eq:Jr_epi_dx}). In the
epicyclic approximation, a $z$ offset only induces an error in $J_z$ --- for
all three orbits the epicyclic approximation is a good description of the
$J_z$ error. An $x$ offset induces an error in $J_r$ and $L_z$. The error in
$J_r$ is somewhat well-described for the thin-disk and thick-disk orbits, and
a poor description for the halo orbit. For $L_z$, the epicyclic approximation
is not a good description for any orbit.}
\label{fig:many_orbit_wrong_ref}
\end{figure*}

@@

\subsection{Velocity Variations} \label{ssec:lsr_var}
We also expect that the local standard of rest (LSR) should vary as a function
of azimuth. We perform this calculation in Appendix~\ref{app:lsr} to estimate
the components of the LSR as a function of azimuth, but performing a best-fit
subtraction to correct for misalignment of the original coordinate system (as
in the previous section) is more involved. Since we find that the variation in
the LSR is less pronounced than for the midplane, and since the dominant
effect is limited to $J_r$ (Figure~\ref{fig:many_orbit_wrong_ref}), we defer
this calculation to future work.

An interesting, qualitative calculation can be made based on the velocity
variations seen in e.g. \citet{2019arXiv190209569F}. The velocity variations
in this work (admittedly found with respect to $L_z$ or radius) are
$\sim5\text{--}10\,\kms$. The vertical frequency of the thin-disk and
thick-disk orbits are $\sim 0.09\,1/\Myr$ and $\sim0.06\,1/\Myr$,
respectively. By dimensional analysis, we therefore expect the midplane
offsets to be of order $\sim 57\text{--}170\,\pc$. We stress that this is a
rough calculation, but it shows that already we see evidence in the data from
velocities for midplane offsets on the order of what we see in both sets of
simulations.

>>>>

Once a Galactocentric coordinate system has been established and a
six-dimensional (6D) phase space measurement of a star has been made, it is
often desirable to convert this measurement into action space to concisely
summarize its projected orbit, model the stellar distribution function, or
find stars with similar dynamical properties. Actions are conserved quantities
that describe the orbit of a star under the assumption of regular, bound
orbits in a system where the equations of motion are separable in a particular
coordinate system. They are the cyclical integral of the canonical momentum
over its conjugate position,
\beq\label{eq:actions}
J_i \equiv
\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,.}
\eeq
Under the assumption of axisymmetry, $i=R,z,\phi$ are the radial, vertical,
and azimuthal coordinates respectively in a cylindrical coordinate system and
the $p_i$ are the conjugate momenta. In a slowly-evolving axisymmetric potential, these
actions are invariant and $J_{\phi} \equiv L_z$
\citep{2008gady.book.....B,2014RvMP...86....1S}.

@@

\begin{deluxetable*}{cCCCCCCCC}
\tablecaption{Relevant dynamical parameters for the $9$ open clusters within
$250\,\pc$ of the Sun, using the relevant initial positions and velocities
from \citet{2018AA...616A..10G}. The neighborhood column is explained in more
detail in Section~\ref{sssec:reconstruction}.\label{tab:real_clusters}}
\tablehead{\colhead{cluster} & \colhead{$N_{\text{memb}}$} & \colhead{distance} & \colhead{$J_R$} &
\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} & \colhead{mass} & \colhead{neighborhood} \\ 
\colhead{ } & \colhead{ } &
\colhead{$\mathrm{pc}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{pc}$} & \colhead{$M_{\odot}$} & \colhead{$\mathrm{pc}$}}
\startdata
alphaPer & 740 & 174.9 & \sim20 & \sim -1700 & \sim0.005 & \sim11 & 352\tablenotemark{b} & 57 \\
Blanco1 & 489 & 237.2 & \sim2 & \sim-1800 & \sim2 & \sim210 & 410\tablenotemark{c} & 1460 \\
ComaBer & 153 & 85.9 & \sim2 & \sim-1900 & \sim0.7 & \sim140 & 112\tablenotemark{d} & 522\\
Hyades & 515 & 47.5 & \sim20 & \sim-1800 & \sim0.3 & \sim80 & 400\tablenotemark{e} & 462 \\
IC2391 & 325 & 151.6 & \sim6 & \sim-1800 & \sim0.02 & \sim20 & \nodata & \nodata \\
IC2602 & 492 & 152.2 & \sim10 & \sim-1700 & \sim0.3 & \sim89 & \nodata & \nodata \\
NGC2451A\tablenotemark{a} & 400 & 193.7 & \sim6 & \sim-1800 & \sim0.2 & \sim73 & \nodata & \nodata \\
Pleiades & 1326 & 135.8 & \sim20 & \sim-1700 & \sim0.4 & \sim96 & 800\tablenotemark{f} & 685 \\
Praesepe & 938 & 186.2 & \sim20 & \sim-1800 & \sim0.6 & \sim120 & 550\tablenotemark{d} & 781
\enddata

@@

We can write down the actions in this approximation as
\citep[Section~3.5.3b][]{2008gady.book.....B},
\beq\label{eq:actions_epi}
\begin{split}
J_{\phi} &= R_g v_c \\
J_R &= \kappa A_R^2 \\
J_z &= \nu A_z^2\text{,}
\end{split}
\eeq
where $R_g$ is the guiding radius of the orbit.

An induced error in $z$ of $\Delta z$ will cause $\Delta A_z \sim \Delta z$.
Similarly, an induced error in $R$ will cause $\Delta A_R \sim \Delta R$.
Making the assumption that $\kappa$ and $\nu$ are constant, we have that
\beq\label{eq:actions_epi_error}
\begin{split}
\Delta J_{\phi} = \Delta R v_c &\implies \frac{\Delta J_{\phi}}{J_{\phi}} = \frac{\Delta R}{R} \\
\Delta J_R = 2 \kappa \Delta A_R &\implies \frac{\Delta J_R}{J_R} = \frac{2\Delta R}{A_R} \\
\Delta J_z = 2 \nu \Delta A_z &\implies \frac{\Delta J_z}{J_z} = \frac{2\Delta z}{A_z}\text{.}
\end{split}
\eeq
Therefore, errors in each action are linearly proportional to the errors in
either the midplane or cylindrical radius. While in detail $\kappa$ and $\nu$
are not constant, we will find that this order of magnitude estimate matches
the numerical estimates quite well.

There is a slight complication in the formula for computing $\Delta J_R/J_R$
and $\Delta J_{\phi}/J_{\phi}$. In a moment we will consider the effect of an inaccuracy
in the distance from the Sun to the Galactic center on actions. However, this
inaccuracy induces an error in $x$, or $\Delta x$. This is not the same as an
inaccuracy in $\Delta R$, but they are related by the following argument. An
error $\Delta x$ will induce an erroneous radius $R_{\text{err}}$ related by
the formula,
\beq
(x+\Delta x)^2 + y^2 = R_{\text{err}}^2\text{.}
\eeq
Keeping only terms to first order in $\Delta x$, we have that,
\beq
\begin{split}
R_{\text{err}}^2 &= R^2 - 2 R \cos{\theta} \Delta x \\
\implies \Delta R &\equiv \abs{R_{\text{err}} - R} = \abs{\cos{\theta}} \Delta x\text{.}
\end{split}
\eeq
Averaging over the circle, we therefore have that
\beq
\avg{\Delta R} = \frac{2}{\pi} \Delta x\text{.}
\eeq
Therefore, the error in $J_R$ and $J_{\phi}$ are related to $\Delta x$ by,
\beq\label{eq:Jr_epi_dx}
\begin{split}
\frac{\Delta J_R}{J_R} &= \frac{4 \Delta x}{\pi A_R} \\
\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{2 \Delta x}{\pi R}\text{.}
\end{split}
\eeq

A similarly rough argument can be made to arrive at the following formulae for
the impact of velocity offsets on the actions,
\beq\label{eq:actions_epi_error}
\begin{split}
\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{\Delta v_{\phi}}{v_c} \\
\frac{\Delta J_R}{J_R} &= \frac{\Delta v_R}{v_{R,\text{max}}} \\
\frac{\Delta J_z}{J_z} &= \frac{\Delta v_z}{v_{z,\text{max}}}\text{,}
\end{split}
\eeq
where $v_{R,\text{max}}$ and $v_{z,\text{max}}$ are the maximum radial and
vertical velocities of the orbit, respectively.

@@

\begin{deluxetable*}{cccccccc}
\tablecaption{Description and names of the three orbits considered in this
work. For the last two columns: $z_{\text{max}}$ the maximum height of the
orbit while $A_R$ is the magnitude of the radial excursions of the orbit,
computed as one half the maximum radius minus the minimum
radius.\label{tab:orbits}}
\tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
velocity} & \colhead{$J_R$} &
\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
\colhead{$A_R$}\\ \colhead{ } &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$}}
\startdata 
thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 & 1.29 \\
thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 & 1.19 \\ 
halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16 & 2.34
\enddata
\end{deluxetable*}

@@

For a systematic offset in $z$ ({\em upper} panels), the
95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
$6.2\,\actunit$ for $J_R$ and $J_z$, respectively. These are fractional errors
of $5.7\,\%$ and $85.7\,\%$, respectively. The error induced in $J_{\phi}$ is
negligible, as expected since $J_{\phi}$ is independent of $z$. It is worth
pointing out that a $100\,\pc$ error in an orbit with
$z_{\text{max}}=850\,\pc$ --- a $12\%$ error
--- induced an $86\%$ error in the computation of $J_z$.

For a systematic offset in $x$ (or Galactic center distance), the
95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$,
$47$, and $0.71\,\actunit$ for $J_R$, $J_{\phi}$ and $J_z$, respectively. These are
fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively.

\begin{figure*}
\plotone{fig/schmactions_one_orbit.pdf}
\caption{The artificial phase-dependence in the observed actions induced by an
error in the Galactocentric coordinate system. We consider here the thick-disk
orbit, which has actions of $(J_R, J_{\phi}, J_z) = (37.9, -1520, 7.0)\,\actunit$
and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We integrate the
orbit according to the procedure laid out in Section~\ref{ssec:action_comp}.
Then, we subtract $100\,\pc$ from the $z$ value ({\em upper panels}) or the
$x$ value ({\em lower panels}) of each position in the orbit, corresponding to
an erroneous observer assuming a midplane ({\em upper}) or solar radius ({\em
lower}) that is off by $100\,\pc$. We then allow our (immortal) observer to
observe the orbit over $1\,\Gyr$ and perform the same orbit integration
procedure at each timestep, and report the values of the actions. The
computation of $J_{\phi}$ is pristine to errors in $z$, with only numerical
artifacts remaining. Only small errors are induced in $J_R$, with the middle
$90\%$ of values over the $\Gyr$ being within $\sim8\%$ of the true $J_R$. As
expected, large errors are induced in $J_z$ with a $100\,\pc$ offset in $z$,
with the middle $90\%$ of values being within $\sim43\%$ of the true $J_z$.
The $x$ offset induces uncertainties in $J_R$ and $J_{\phi}$ of $\sim21\%$ and
$\sim3\%$, respectively. A $\sim3\%$ error in $J_z$ is also induced.}
\label{fig:one_orbit_wrong_ref}
\end{figure*}

\begin{figure}
\plotone{fig/schmactions_Jz_hist.pdf}
\caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
values have a bias (from e.g. an incorrectly computed midplane), then the
induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
sort of error propagation must take this into account. Intuition for the shape
of each panel is given in the text. We also plot one half the 95\uth{}
percentile minus the 5\uth{} percentile of each distribution as a horizontal
arrow anchored on the true $J_z$ value. We call this $\Delta J_z$ and will use
it (along with the similarly defined $\Delta J_R$ and $\Delta J_{\phi}$) to
empirically describe the error distribution. We see that $\Delta J_z$ roughly
corresponds to the distance from the true $J_z$ value to one of the modes of
the error distribution.}
\label{fig:Jz_hist}
\end{figure}

@@

The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
in each action for an offset in the $z$-component (i.e. the Galactic
midplane). In the {\em bottom row} we consider offsets in the $x$ component
(i.e. the Solar radius). The {\em left}, {\em middle}, and {\em right} columns
show the fractional spread in the values computed for $J_z$, $J_{\phi}$, and $J_R$,
respectively. We compute the fractional spread $\Delta J_i/J_i$ as one half
the 95\uth{} minus 5\uth{} percentile of the action computed using the offset
coordinate system over the course of the first $\Gyr$, divided by the true
action (computed in the correct coordinate system).

In the {\em upper middle} panel of Figure~\ref{fig:many_orbit_wrong_ref},
there is essentially no spread in the determination of $J_{\phi}$. This is expected
since, in principle, $J_{\phi}$ only depends on the $x$- and $y$-components of the
position and velocity of the stars\footnote{In practice, however, $J_{\phi}$ is
computed as part of the torus-fitting method.}, and is thus unaffected by
offsets in $z$. Indeed, the result we found in
Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all
orbit types.

@@

For the offset in the Solar radius ({\em bottom panels}), the error is largest
for $J_R$, with some deviations resulting in $J_{\phi}$ and relatively small
deviations in $J_z$. In the {\em bottom middle} and {\em bottom right} panels
all three lines nearly overlap.

\begin{figure*}
\plotone{fig/schmactions_many_orbits.pdf}
\caption{We report one half the 95\uth{} minus 5\uth{} percentile of the error
in the measured action ($\Delta J_i$) from coordinate system errors for the
thin, thick, and halo orbits (Table~\ref{tab:orbits}). See discussion in the
text for the justification in using this to measure the magnitude of the
induced error. The {\em left}, {\em center}, and {\em right} panels show the
result for $J_R$, $J_{\phi}$, and $J_z$, respectively. The {\em upper} panels
consider an offset in $z$ and the {\em lower} panels consider an offset in $x$
(equivalently, an offset in the Solar radius). In some panels we also plot as
dashed lines the epicyclic prediction of the induced action error
(Equations~\ref{eq:actions_epi_error}~and~\ref{eq:Jr_epi_dx}). In the
epicyclic approximation, a $z$ offset only induces an error in $J_z$ --- for
all three orbits the epicyclic approximation is a good description of the
$J_z$ error. An $x$ offset induces an error in $J_R$ and $J_{\phi}$. The error in
$J_R$ is somewhat well-described for the thin-disk and thick-disk orbits, and
a poor description for the halo orbit. For $J_{\phi}$, the epicyclic approximation
is not a good description for any orbit.}
\label{fig:many_orbit_wrong_ref}
\end{figure*}

@@

\subsection{Velocity Variations} \label{ssec:lsr_var}
We also expect that the local standard of rest (LSR) should vary as a function
of azimuth. We perform this calculation in Appendix~\ref{app:lsr} to estimate
the components of the LSR as a function of azimuth, but performing a best-fit
subtraction to correct for misalignment of the original coordinate system (as
in the previous section) is more involved. Since we find that the variation in
the LSR is less pronounced than for the midplane, and since the dominant
effect is limited to $J_R$ (Figure~\ref{fig:many_orbit_wrong_ref}), we defer
this calculation to future work.

An interesting, qualitative calculation can be made based on the velocity
variations seen in e.g. \citet{2019arXiv190209569F}. The velocity variations
in this work (admittedly found with respect to $J_{\phi}$ or radius) are
$\sim5\text{--}10\,\kms$. The vertical frequency of the thin-disk and
thick-disk orbits are $\sim 0.09\,1/\Myr$ and $\sim0.06\,1/\Myr$,
respectively. By dimensional analysis, we therefore expect the midplane
offsets to be of order $\sim 57\text{--}170\,\pc$. We stress that this is a
rough calculation, but it shows that already we see evidence in the data from
velocities for midplane offsets on the order of what we see in both sets of
simulations.
